### 2023-02-16 notebooks/conformal.qmd
woohoo

```{julia}
using CCE
using ConformalPrediction
using CounterfactualExplanations
using CounterfactualExplanations.Data
using Flux
using MLJBase
using MLJFlux
using Plots
```

```{julia}
counterfactual_data = load_linearly_separable()
```

```{julia}
builder = MLJFlux.@builder Chain(
    Dense(n_in, 32, relu),
    Dense(32, n_out)
)
clf = NeuralNetworkClassifier(builder=builder, epochs=100)
```

```{julia}
X = table(permutedims(counterfactual_data.X))
y =  counterfactual_data.output_encoder.labels
conf_model = conformal_model(clf; method=:simple_inductive)
mach = machine(conf_model, X, y)
fit!(mach)
```

## Counterfactual Explanation

```{julia}
M = CCE.ConformalModel(conf_model, mach.fitresult)
```

```{julia}
x = select_factual(counterfactual_data,rand(1:size(counterfactual_data.X,2)))
y_factual = predict_label(M, counterfactual_data, x)[1]
target = counterfactual_data.y_levels[counterfactual_data.y_levels .!= y_factual][1]
```

```{julia}
p1 = contourf(mach.model, mach.fitresult, X, y; plot_classification_loss=true, target=target, zoom=0)
p2 = contourf(mach.model, mach.fitresult, X, y; plot_set_loss=true, zoom=0)
plot(p1, p2, size=(800,320))
```

```{julia}
#| output: true
#| echo: false

ordered_names = [
    "Generic (γ=0.5)",
    "Generic (γ=0.9)",
    "Conformal (λ₂=1)",
    "Conformal (λ₂=10)"
]

# Generators:
generators = Dict(
    ordered_names[1] => GenericGenerator(opt = opt, decision_threshold=0.5),
    ordered_names[2] => GenericGenerator(opt = opt, decision_threshold=0.9),
    ordered_names[3] => CCE.ConformalGenerator(opt=opt, λ=[0.1,1]),
    ordered_names[4] => CCE.ConformalGenerator(opt=opt, λ=[0.1,10]),
)

counterfactuals = Dict([name => generate_counterfactual(x, target, counterfactual_data, M, gen; initialization=:identity) for (name, gen) in generators])

# Plots:
plts = []
for name ∈ ordered_names
    ce = counterfactuals[name]
    plt = plot(ce; title=name, colorbar=false, ticks = false, legend=false, zoom=0)
    plts = vcat(plts..., plt)
end
_n = length(generators)
img_size = 300
plot(plts..., size=(_n * img_size,1.05*img_size), layout=(1,_n))
```