### 2023-04-04 notebooks/synthetic.qmd
uh

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks);
```

# Synthetic data

```{julia}
#| output: false

# Data:
datasets = Dict(
    :linearly_separable => load_linearly_separable(),
    :overlapping => load_overlapping(),
    :moons => load_moons(),
    :circles => load_circles(),
    :multi_class => load_multi_class(),
)

# Hyperparameters:
cvgs = [0.5, 0.75, 0.90, 0.95, 0.99]
temps = [0.05, 0.1, 0.5, 1.0, 2.0]
Λ = [0.1, 1.0, 5.0]

# Classifiers:
epochs = 100
link_fun = sigmoid
logreg = NeuralNetworkClassifier(builder=MLJFlux.Linear(σ=link_fun), epochs=epochs)
mlp = NeuralNetworkClassifier(builder=MLJFlux.MLP(hidden=(32,), σ=link_fun), epochs=epochs)
ensmbl = EnsembleModel(model=mlp, n=5)
classifiers = Dict(
    # :logreg => logreg,
    :mlp => mlp,
    # :ensmbl => ensmbl,
)

# Search parameters:
target = 2
factual = 1
```

```{julia}
results = DataFrame()
for (dataname, data) in datasets

    # Data:
    X = table(permutedims(data.X))
    y = data.output_encoder.labels

    for (clf_name, clf) in classifiers, cov in cvgs

        # Classifier and coverage:
        conf_model = conformal_model(clf; method=:simple_inductive, coverage=cov)
        mach = machine(conf_model, X, y)
        fit!(mach)
        M = CCE.ConformalModel(mach.model, mach.fitresult)

        # Set up CCE:
        yhat = predict_label(M, data)
        factual_label = data.y_levels[factual]
        target_label = data.y_levels[target]
        x = select_factual(data,rand(findall(yhat .== factual_label)))

        for λ in Λ, temp in temps

            # CCE for given classifier, coverage, temperature and λ:
            generator = CCEGenerator(temp=temp, λ=λ)
            @assert predict_label(M, data, x) != target_label
            ce = try
                generate_counterfactual(
                    x, target_label, data, M, generator;
                    initialization=:identity,
                    converge_when=:generator_conditions,
                )
            catch
                missing
            end

            _results = DataFrame(
                dataset = dataname,
                classifier = clf_name,
                coverage = cov,
                temperature = temp,
                λ = λ,
                counterfactual = ce,
                factual = factual_label,
                target = target_label,
            )
            append!(results, _results)

        end

    end

end
```

```{julia}

```