### 2023-04-25 notebooks/eccco.qmd
work on paper

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks)
```