### 2023-05-02 notebooks/real_world.qmd

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks)
```

# Real-World Data

```{julia}
# Hyper:
_retrain = false
_regen = false

# Data:
n_obs = 10000
datasets = load_tabular_data(n_obs; drop=:credit_default)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

First, let's create a couple of image classifier architectures:

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 64
activation = Flux.relu
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden, activation),
    Dense(n_hidden, n_out),
)
n_ens = 5                                   # number of models in ensemble
_loss = Flux.Losses.logitcrossentropy       # loss function
_finaliser = x -> x                         # finaliser function
```

```{julia}
# JEM parameters:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,1e-2]      # penalty strengths
```

```{julia}
# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=builder, 
    epochs=epochs,
    batch_size=batch_size,
    finaliser=_finaliser,
    loss=_loss,
)

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=n_ens)

# Joint Energy Model:
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    epochs=epochs,
    batch_size=batch_size,
    finaliser=_finaliser,
    loss=_loss,
    jem_training_params=(
        α=α,verbosity=10,
    ),
    sampling_steps=20,
)

# JEM with adversarial training:
jem_adv = deepcopy(jem)
# jem_adv.adv_training = true

# Deep Ensemble of Joint Energy Models:
jem_ens = EnsembleModel(model=jem, n=n_ens)

# Deep Ensemble of Joint Energy Models with adversarial training:
# jem_ens_plus = EnsembleModel(model=jem_adv, n=n_ens)

# Dictionary of models:
models = Dict(
    "MLP" => mlp,
    "MLP Ensemble" => mlp_ens,
    "JEM" => jem,
    "JEM Ensemble" => jem_ens,
    # "JEM Ensemble+" => jem_ens_plus,
)
```

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(mod; mod_name=mod_name) for (mod_name, mod) in models)
    Serialization.serialize(joinpath(output_path,"mnist_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"mnist_models.jls"))
end
```

```{julia}
# Plot generated samples:
n_regen = 150
if _regen 
    for (mod_name, mod) in model_dict
        if ECCCo._has_sampler(mod)
            sampler = ECCCo._get_sampler(mod)
        else
            K = length(counterfactual_data.y_levels)
            input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
            𝒟x = Uniform(extrema(counterfactual_data.X)...)
            𝒟y = Categorical(ones(K) ./ K)
            sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
        end
        opt = ImproperSGLD()
        f(x) = logits(mod, x)

        _w = 1500
        plts = []
        neach = 10
        for i in 1:10
            x = sampler(f, opt; niter=n_regen, n_samples=neach, y=i)
            plts_i = []
            for j in 1:size(x, 2)
                xj = x[:,j]
                xj = reshape(xj, (n_digits, n_digits))
                plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
            end
            plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
            plts = [plts..., plt]
        end
        plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1), plot_title=mod_name)
        savefig(plt, joinpath(output_images_path, "mnist_generated_$(mod_name).png"))
        display(plt)
    end
end
```

```{julia}
# Evaluate models:

measure = Dict(
    :f1score => multiclass_f1score, 
    :acc => accuracy, 
    :precision => multiclass_precision
)
model_performance = DataFrame()
for (mod_name, mod) in model_dict
    # Test performance:
    test_data = load_mnist_test()
    _perf = CounterfactualExplanations.Models.model_evaluation(mod, test_data, measure=collect(values(measure)))
    _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
    _perf.mod_name .= mod_name
    model_performance = vcat(model_performance, _perf)
end
Serialization.serialize(joinpath(output_path,"mnist_model_performance.jls"), model_performance)
CSV.write(joinpath(output_path, "mnist_model_performance.csv"), model_performance)
model_performance
```

### Different Models

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.5,0.1,0.5],
    temp=0.1,η=0.01,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = Flux.Optimise.Adam(η),
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
)

    # Setup:
    Random.seed!(rng)
    if x isa Int
        x = reshape(counterfactual_data.X[:,rand(findall(labels.==x))],input_dim,1)
    end

    # Generate counterfactuals using ECCCo generator:
    eccco_generator = ECCCoGenerator(
        λ=λ, 
        temp=temp, 
        opt=opt,
    )

    ces = Dict()
    for (mod_name, mod) in model_dict
        ce = generate_counterfactual(
            x, target, counterfactual_data, mod, eccco_generator; 
            decision_threshold=γ, max_iter=T,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        ces[mod_name] = ce
    end
    _plt_order = map(x -> findall(collect(keys(model_dict)) .== x)[1], plt_order)

    # Plot:
    p1 = Plots.plot(
        convert2image(MNIST, reshape(x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title="Factual"
    )

    plts = []
    for (_name,ce) in ces
        _x = CounterfactualExplanations.counterfactual(ce)
        _phat = target_probs(ce)
        _title = "$_name (p̂=$(round(_phat[1]; digits=3)))"
        plt = Plots.plot(
            convert2image(MNIST, reshape(_x,28,28)),
            axis=nothing, 
            size=(img_height, img_height),
            title=_title
        )
        plts = [plts..., plt]
    end
    plts = plts[_plt_order]
    plts = [p1, plts...]
    plt = Plots.plot(plts...; size=(img_height*length(plts),img_height), layout=(1,length(plts)))

    return plt, eccco_generator
end
```

```{julia}
plt, eccco_generator = _plot_eccco_mnist()
display(plt)
savefig(plt, joinpath(output_images_path, "mnist_eccco.png"))
```

### All digits

```{julia}
function plot_mnist(
    factual::Int,target::Int;
    generator::AbstractGenerator,
    model::AbstractFittedModel=model_dict["JEM Ensemble"],
    data::CounterfactualData=counterfactual_data,
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
    _plot_title::Bool=true,
    kwargs...,
)
    decision_threshold = !isdefined(kwargs, :decision_threshold) ? 0.9 : decision_threshold
    max_iter = !isdefined(kwargs, :max_iter) ? 100 : max_iter
    initialization = !isdefined(kwargs, :initialization) ? :identity : initialization
    converge_when = !isdefined(kwargs, :converge_when) ? :generator_conditions : converge_when

    x = reshape(data.X[:,rand(findall(predict_label(model, data).==factual))],input_dim,1)
    ce = generate_counterfactual(
        x, target, data, model, generator; 
        decision_threshold=decision_threshold, max_iter=max_iter,
        initialization=initialization,
        converge_when=converge_when,
        kwargs...
    )

    _title = _plot_title ? "$(factual) -> $(target)" : ""

    _x = CounterfactualExplanations.counterfactual(ce)
    plt = Plots.plot(
        convert2image(MNIST, reshape(_x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    return plt
end
```

```{julia}
if _regen
    function plot_all_digits(rng=1;verbose=true,kwargs...)
        plts = []
        for i in 0:9
            for j in 0:9
                @info "Generating counterfactual for $(i) -> $(j)"
                plt = plot_mnist(i,j;kwargs...,rng=rng)
                !verbose || display(plt)
                plts = [plts..., plt]
            end
        end
        plt = Plots.plot(plts...; size=(img_height*10,img_height*10), layout=(10,10))
        return plt
    end
    plt = plot_all_digits(generator=eccco_generator)
    savefig(plt, joinpath(output_images_path, "mnist_eccco_all_digits.png"))
end
```

### Different Generators

```{julia}
# Setup:
model = model_dict["JEM Ensemble"]

# Benchmark generators:
generator_dict = Dict(
    :wachter => generic_generator,
    :revise => revise_generator,
    :greedy => greedy_generator,
    :eccco => eccco_generator,
)

ces = Dict()
for (gen_name, gen) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, model, gen; 
        decision_threshold=γ, max_iter=T,
        initialization=:identity,
        converge_when=:generator_conditions,
    )
    ces[gen_name] = ce
end
plt_order = sortperm(collect(keys(ces)))

# Plot:
p1 = Plots.plot(
    convert2image(MNIST, reshape(x_factual,28,28)),
    axis=nothing, 
    size=(img_height, img_height),
    title="Factual"
)

plts = []
for (_name,ce) in ces
    _x = CounterfactualExplanations.counterfactual(ce)
    _phat = target_probs(ce)
    _title = "$_name (p̂=$(round(_phat[1]; digits=3)))"
    plt = Plots.plot(
        convert2image(MNIST, reshape(_x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    plts = [plts..., plt]
end
plts = plts[plt_order]
plts = [p1, plts...]
plt = Plots.plot(plts...; size=(img_height*length(plts),img_height), layout=(1,length(plts)))
display(plt)
savefig(plt, joinpath(output_images_path, "mnist_all_generators.png"))
```

## Benchmark

```{julia}
# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.Evaluation.validity,
    CounterfactualExplanations.Evaluation.redundancy,
]

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="MNIST",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
)
CSV.write(joinpath(output_path, "mnist_benchmark.csv"), bmk())
```

```{julia}
@chain bmk() begin
    @group_by(dataname, generator, model, variable)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable == "distance_from_energy")
end
```

```{julia}
df = @chain bmk() begin
    @filter(variable in [
        "distance_from_energy",
        "distance_from_targets",
        "distance",])
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
    @mutate(variable = ifelse.(variable .== "distance", "Cost", variable))
end
plt = AlgebraOfGraphics.data(df) * visual(BoxPlot) * 
    mapping(:generator, :value, row=:variable, col=:model, color=:generator)
plt = draw(
    plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=150, height=120), 
    facet=(; linkyaxes=:minimal)
)   
display(plt)
save(joinpath(output_images_path, "mnist_benchmark.png"), plt, px_per_unit=5)
```