### 2023-05-07 notebooks/results.qmd
work on table

```{julia}
include("notebooks/setup.jl")
eval(setup_notebooks)
```

# Results

```{julia}
df = DataFrame()
for _file in filter(endswith("_benchmark.csv"), readdir(output_path)) 
    x = joinpath(output_path, _file)
    _df = CSV.read(x, DataFrame)
    df = vcat(df, _df, cols=:intersect)
end
synth = ("Moons", "Circles")
df.source .= ifelse.(df.dataname .∈ [synth], :synthetic, :real) 
```

```{julia}
tab = @chain df begin
    @group_by(dataname, generator, model, variable, source)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable ∈ ["distance_from_energy", "distance_from_targets"])
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
end
```