### 2023-08-09 dev/rebuttal/support.qmd
nearly there

---
format: 
    pdf:
        fontsize: 10pt
---

![Turning a 6 into a 0. Top row: ECCCo. Bottom row: Wachter. The different underlying models across columns: (a) MLP, (b) Small Ensemble $n=5$, (c) Large Ensemble $n=50$, (d) LeNet-5, (e) JEM, (f) JEM Ensemble.](www/mnist_6to0_7.png){#fig-mnist-1 width="80%"}

![Turning a 9 into a 7. Rows and columns as in @fig-mnist-1.](www/mnist_9to7_1.png){#fig-mnist-2 width="80%"}

<!-- ![Turning a 4 into a 1. Top row: ECCCo. Bottom row: Wachter. The different underlying models across columns: (a) MLP, (b) Small Ensemble $n=5$, (c) Large Ensemble $n=50$, (d) LeNet-5, (e) JEM, (f) JEM Ensemble.](www/mnist_4to1_11.png){#fig-mnist-3 width="75%"} -->