### 2023-08-10 notebooks/fmnist.qmd

```{julia}
include("$(pwd())/notebooks/setup.jl")
eval(setup_notebooks)
```

# FashionMNIST

## Anecdotal Evidence

### Examples in Introduction

#### Wachter and JSMA

```{julia}
Random.seed!(2023)

# Data:
counterfactual_data = load_fashion_mnist()
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
input_dim, n_obs = size(counterfactual_data.X)
M = load_fashion_mnist_mlp()

# Target:
factual_label = 9
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 7
factual = predict_label(M, counterfactual_data, x_factual)[1]
γ = 0.9

# Training params:
T = 100
```

### ECCCo

```{julia}
function pre_process(x; noise::Float32=0.03f0)
    ϵ = Float32.(randn(size(x)) * noise)
    x += ϵ
    return x
end
```

```{julia}
# Hyper:
_retrain = false
_regen = false

# Data:
n_obs = 10000
counterfactual_data = load_fashion_mnist(n_obs)
counterfactual_data.X = pre_process.(counterfactual_data.X)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

First, let's create a couple of image classifier architectures:

```{julia}
# Model parameters:
epochs = 10
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 512
activation = Flux.relu
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden, activation),
    # Dense(n_hidden, n_hidden, activation),
    # Dense(n_hidden, n_hidden, activation),
    Dense(n_hidden, n_out),
)
n_ens = 5                                  # number of models in ensemble
_loss = Flux.Losses.logitcrossentropy       # loss function
_finaliser = x -> x                         # finaliser function
```

```{julia}
# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=builder, 
    epochs=epochs,
    batch_size=batch_size,
    finaliser=_finaliser,
    loss=_loss,
)

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=n_ens)

# Dictionary of models:
models = Dict(
    # "MLP" => mlp,
    "Ensemble" => mlp_ens,
)

Serialization.serialize(joinpath(output_path,"fashion_mnist_architectures.jls"), models)
```

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(mod; mod_name=mod_name) for (mod_name, mod) in models)
    Serialization.serialize(joinpath(output_path,"fashion_mnist_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"fashion_mnist_models.jls"))
end
```

```{julia}
# Plot generated samples:
n_regen = 500
if _regen 
    for (mod_name, mod) in model_dict
        K = length(counterfactual_data.y_levels)
        input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
        𝒟x = Uniform(extrema(counterfactual_data.X)...)
        𝒟y = Categorical(ones(K) ./ K)
        sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size, prob_buffer=0.0)
        opt = ImproperSGLD()
        f(x) = logits(mod, x)

        _w = 1000
        plts = []
        neach = 1
        for i in 1:10
            x = sampler(f, opt; niter=n_regen, n_samples=neach, y=i)
            plts_i = []
            for j in 1:size(x, 2)
                xj = x[:,j]
                xj = reshape(xj, (n_digits, n_digits))
                plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
            end
            plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
            plts = [plts..., plt]
        end
        plt = Plots.plot(plts..., size=(_w,_w/10), layout=(1,10), plot_title=mod_name)
        savefig(plt, joinpath(output_images_path, "fashion_mnist_generated_$(mod_name).png"))
        display(plt)
    end
end
```

```{julia}
# Evaluate models:

measure = Dict(
    :f1score => multiclass_f1score, 
    :acc => accuracy, 
    :precision => multiclass_precision
)
model_performance = DataFrame()
for (mod_name, mod) in model_dict
    # Test performance:
    test_data = load_fashion_mnist_test()
    _perf = CounterfactualExplanations.Models.model_evaluation(mod, test_data, measure=collect(values(measure)))
    _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
    _perf.mod_name .= mod_name
    model_performance = vcat(model_performance, _perf)
end
Serialization.serialize(joinpath(output_path,"fashion_mnist_model_performance.jls"), model_performance)
CSV.write(joinpath(output_path, "fashion_mnist_model_performance.csv"), model_performance)
model_performance
```

```{julia}
label_names = [
    "T-shirt/top",
    "Trouser",
    "Pullover",
    "Dress",
    "Coat",
    "Sandal",
    "Shirt",
    "Sneaker",
    "Bag",
    "Boot",
]

function plot_fmnist(
    x::Int=3, targets::Vector{Int}=[1];
    generator = ECCCoGenerator(),
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
    T::Int = 100,
    use_class_loss::Bool = true,
    model=model_dict["Ensemble"],
    img_height::Int = img_height,
    test_data::Bool = false,
    init::Symbol = :identity,
    noise::Float32 = 0.0f0,
    kwrgs...,
)

    # Setup:
    Random.seed!(rng)
    x_fact = counterfactual_data.X[:,rand(findall(labels.==x))][:,:]
    x_fact = pre_process(x_fact, noise=noise)

    if test_data
        data = load_mnist_test()
    else
        data = counterfactual_data
    end

    # Plot:
    p1 = Plots.plot(
        convert2image(MNIST, reshape(x_fact,28,28)),
        axis=([], false),
        size=(img_height, img_height),
        title="$(label_names[x+1]) (factual)",
        dpi=300
    )

    plts = [p1]

    for t in targets
        ce = generate_counterfactual(
            x_fact, t, data, model, generator; 
            decision_threshold=γ, max_iter=T,
            initialization=init,
            converge_when=:generator_conditions,
        )
        _x = CounterfactualExplanations.counterfactual(ce)
        plt = Plots.plot(
            convert2image(MNIST, reshape(_x,28,28)),
            axis=([], false), 
            size=(img_height, img_height),
            title="$(label_names[t+1])", 
            dpi=300
        )
        plts = [plts..., plt]
    end

    plt = Plots.plot(plts..., size=(img_height*length(plts),img_height), layout=(1,length(plts)), kwrgs...)

    return plt
end

```

```{julia}
rng = 42
gen = ECCCoGenerator(
    λ=[0.1,0.30,0.30], 
    nsamples=25,
    nmin=10,
)
plts = []

plt = plot_fmnist(3, [1]; rng=rng, generator=gen, img_height=200)
display(plt)
push!(plts, plt)
savefig(plt, "dev/rebuttal/www/fmnist_dress.png")

plt = plot_fmnist(9, [5]; rng=rng, generator=gen, img_height=200)
display(plt)
push!(plts, plt)
savefig(plt, "dev/rebuttal/www/fmnist_boot.png")

plt = plot_fmnist(2, [4]; rng=rng, generator=gen, img_height=200)
display(plt)
push!(plts, plt)
savefig(plt, "dev/rebuttal/www/fmnist_pullover.png")
```