### 2023-11-06 experiments/notebooks/plots.qmd
plot reviewer 1

---
format: pdf
---

# Plots

```{julia}
using Pkg; Pkg.activate("experiments")
include("$(pwd())/experiments/setup_env.jl")
```

## Counterfactual Path - MNIST

```{julia}
outcome = Serialization.deserialize("results/moons_outcome.jls")
data = outcome.exper.counterfactual_data
```

```{julia}
factual = data.y_levels[2]
target = data.y_levels[1]
```

```{julia}
n_samp = 100
n_rand = 500
alpha = 0.2
plts = []
for (mod_name, model) in outcome.model_dict
    t = get_target_index(data.y_levels, target)
    E(x) = -logits(model, x)[t, :]
    x_samp = data.X[:,rand(findall(data.output_encoder.labels.==target),n_samp)]
    x_rand = Float32.(randn(size(data.X,1), n_rand))
    dist = map(eachcol(x_rand)) do x
        mean(map(y -> norm(x.-y),eachcol(x_samp)))
    end
    plt = scatter(
        E(x_rand), dist;
        fillalpha=alpha, label="", title=mod_name, smooth=:true,
        lc=:red, lw=2
    )
    push!(plts, plt)
end
plot(plts..., layout=(1,length(outcome.model_dict)), size=(1000,250))
```