### 2023-12-19 paper/contents/table_gen_params.tex

\begin{table}

\caption{Generator hyperparameters: the optimiser step size ($\eta$); penalty strengths where $\lambda_1$ applies to all generators but \textit{Schut} and the other parameter are specific to \textit{ECCCo}; finally, the strength for the Ridge penalty on energy for \textit{ECCCo}.\label{tab:genparams} \newline}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{rrrrrr}
\toprule
Dataset & $\eta$ & $\lambda_1$ & $\lambda_2$ & $\lambda_3$ & Ridge penalty\\
\midrule
Linearly Separable & 0.05 & 0.10 & 0.1 & 0.5 & 0.0\\
Moons & 0.05 & 0.10 & 0.1 & 0.2 & 0.0\\
Circles & 0.05 & 0.10 & 0.1 & 0.2 & 0.0\\
California Housing & 0.05 & 0.10 & 0.1 & 0.5 & 0.0\\
GMSC & 0.05 & 0.10 & 0.1 & 0.5 & 0.0\\
German Credit & 0.05 & 0.10 & 0.1 & 0.1 & 0.5\\
MNIST & 0.05 & 0.01 & 0.1 & 0.3 & 0.0\\
Fashion MNIST & 0.05 & 0.01 & 0.1 & 0.3 & 0.0\\
\bottomrule
\end{tabular}
\end{table}